\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2018-02-22 redaction/dhp_july2017.tex
Changement texte, encore et toujours

Trade costs are central to international economic analysis. In particular, they are considered a major obstacle to international economic integration and international trade flows. According to \cite{Jacks08}, the trade cost decline explains around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse (but see \cite{corlay_2017}). After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO after 1995) multilateral agreements. Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000. Several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} and \citealp{Disdier_Head08}). Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.

Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se} (\citet{daudin_logistique_2003, daudin_transactions_2005}). In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs would represent a 21\% markup over production costs. Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) indicates that the impact of transport costs on trade flows is not negligible. If much of trade policy barriers have been removed over the second half of the 20th century, this suggests that the transport costs component remains large and deserves attention. International transport costs accordingly stand at the heart of the paper.\smallskip

@@

Our results may be summarized in two main new findings. First, additive transport costs are quantitatively sizeable. On average over 1974-2013, the additive cost is estimated to be 1.8\% and 2.9\% of the export price in air and vessel transport respectively. This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period). Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport mode. To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs. This represents a valuable insight for calibrating related international trade or business cycle models. We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs. Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor~2. Additionally, the goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degrees of freedom. \textbf{attention maintenant en annexe. Le dire en intro, est ce pertinent?}

@@

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian firm-level export data in 2004, they find that the ratio of additive to ad-valorem trade costs is about 14\% of the export price. Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs. Yet, our paper complements their findings in two main respects. First, while our study covers a narrower set of trade costs focused on transport costs, their data only allows recovering the ratio of additive to ad-valorem costs as a fraction of export price; in contrast, our empirical framework provides a distinct measure of each additive and ad-valorem cost component. From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect. Based on the estimated share of transport costs in total trade costs provided by \citet{anderson_wincoop_jel} (around 15\%), straightforward calculation yields a ratio of additive to multiplicative transport (not trade) costs of 2.1\% (14\% $\times$ 15\%) in \citet{Irrazabal_2015}. This falls in the range of our own estimates, making us confident in our quantitative assessment of the size of international transport costs. \footnote{See Section \ref{sec:results_decomposition} for more details.} Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} and \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

\textbf{enleve ancienne footnote plus haut qui reprenait cette difference, a mettre en footnote ici? en commentaire en dessous}
%\footnote{As we discuss below, the closest paper to us in this respect is \citet{Irrazabal_2015}. However, given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$). By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.}

@@

\textbf{stop here}

@@

Under Model (\textit{(a)}), the estimated equation simplifies as:\footnote{One may also object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs. This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$. The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results. These results are reported in the Online Appendix, available on the authors' webpages.}

@@

After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component, $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode. When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is} = \widehat{\tau}_{i} \times \widehat{\tau}_{s}$.\footnote{In this case, notice that the equation could be estimated relying on a linear form. To preserve comparability of the results, we keep the same non-linear estimation method in both cases though.} Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme. We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{ice}$, \textit{b)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode. These results are reported in Section \ref{sec:results_decomposition}.

@@

\begin{table}[htbp]
  \centering
  \footnotesize{
  \caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    \multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
    Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
    \multicolumn{5}{l}{\textbf{Model (a) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice}$, in \%)}  \\ \hline
    Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
   \textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} & & & & \\ \hline
    Mean  & 3.2 & 2.5 & 3.3 & 2.4 \\
    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \textit{Additive term }& & & &   \\ \hline
    \textit{In \% of the export price} &&&& \\ \hline
        Mean  & 2.9 & 1.8 & 2.8 & 1.9 \\
    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
   \textit{ In USD per kg traded }&&&& \\ \hline
    Mean & 0.09	&1.12	&0.08&	2.16 \\
    Median & 0.07	&1.04	&0.06&	1.06 \\ \hline
    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
    \multicolumn{5}{l}{\textbf{Data}  } \\ \hline
    Transport costs ($p/\widetilde{p}$, in \%) & & & & \\ \hline
    Mean & 5.3 & 5.0& 5.6&3.9 \\
    Median & 4.3 & 2.0 & 4.4& 1.9 \\ \hline
    Export price ($\widetilde{p}$), in USD & & & & \\
    Mean & 16.0 &	6488.4	&9.6	&6643.6 \\
    Median & 4.1	& 142.5	& 4.1	& 142.2 \\ \hline
    \# obs. & 29279 & 28207 & 29317 & 27680 \\
    \# origin country & 188 & 191 & 188 & 189 \\
    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
  \end{tabular}
    \end{center}}
\parbox[l]{10cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value relative to total trade flows. The additive term is expressed in fraction of fas price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}\\
\parbox[l]{10cm}{\tiny{Notes: For the  4-digit classification, statistics for observed data have been calculated for the same set of years as used for estimation, i.e. 1974, 1981, 1989, 2001, 2009, 2013.}}
\end{table}%

@@

Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs. On average over the whole period, additive costs are estimated to around 10 cents and 1.12 USD per kilogram exported, in maritime and air transport cost respectively (at the 3-digit level), indicating that freight rates are much higher in air transport. Expressed in terms of the export price, the additive costs amount to 2.9\% and 1.8 \% for ocean shipping and air transport respectively. This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\% and 1.8\% for vessel and air shipping respectively (on average over the period). Put it differently, the additive component amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air. This shows that the per-kg cost dimension is quantitatively sizeable. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly a factor 2. Further, these results provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for the related more theoretical papers, which need to calibrate their models.\smallskip

Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question. Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that the ratio of additive to ad-valorem trade costs amounts to 14\% of the export price, ie $\frac{t/\tau}{\widetilde{p}}=14\%$ expressed in our terminology. One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem). From this, we can rebuilt the ratio in similar terms to them -- allowing us to gain in generality in this respect. Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.8\% in air transport. This may sound surprisingly low, in contrast to the 14\% obtained by \citet{Irrazabal_2015}. This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case. While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs (as we start from the gap between the import and the export prices). From \citet{anderson_wincoop_jel}, we can infer that monetary transport costs approximatively represent 15\% of total trade costs.\footnote{\citet{anderson_wincoop_jel} estimate that monetary transport costs represent a 11\% markup over production costs, and that total international trade costs represent 74\%, so transport costs amount to 11/74= 15\% of total trade costs.} Plugging this into \citet{Irrazabal_2015} results, allows to deduce a ratio $t/\tau$ of international transport costs to the export price $\widetilde{p}$ equal to 2.1\% ($0.15\times 0.14$). This stands in line with our own results, and makes us confident in our quantitative assessment of the size of the additive component in international transport costs.\footnote{See also Appendix \ref{app:diagnostic_test}, which confirms the importance of the additive cost component through another perspective angle based on quality-of-fit diagnostic tests.}\smallskip

@@

\paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}. He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f$ the shipping charge per kg shipped. Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p_{ikt}}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, the transport cost measure writes down as:

@@

We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs. To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price, ie $\beta$, according to:
\begin{eqnarray*}
\beta_{ikt} &\equiv& \frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}}
\end{eqnarray*}

@@

The above results point out that the share of additive costs in total transport costs is not constant across the triplet year/product/origin country, thereby contrasting the implicit Hummel's (2007) assumption. This drives us to provide a decomposition of the transport costs trends (between the trade composition effects and the ``pure'' transport costs time trends), that explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} One may speculate that making the (empirically relevant) assumption of a varying share of additive costs over time, country and product is not anecdotal. Importantly, we confirm the key role of this assumption in analyzing the underlying determinants of transport costs trend patterns. We start presenting our estimation strategy.\smallskip

>>>>

Trade costs are central to international economic analysis. In particular, they are considered a major obstacle to international economic integration and international trade flows. After the Second World War though, the reduction of policy trade costs have been the workhorse of the GATT (WTO after 1995) rounds of multilateral negotiations. Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000. Despite this, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\cite{Head_Mayer04} or \cite{Disdier_Head08}, to name a few). Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.

Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}. In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs would represent a 21\% markup over production costs. Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) indicates that the impact of transport costs on trade flows is not negligible. If much of trade policy barriers have been removed over the second half of the 20th century, this suggests that the transport costs component remains large and deserves attention. International transport costs accordingly stand at the heart of the paper.\smallskip

\textbf{j'ai enleve les references a Guillaume parce que ca me semble pas bien place ici, je ne sais pas bien ce que ca veut dire a cet endroit. On peut rediscuter de ou les mettre.}

@@

Our results may be summarized in two main new findings. First, additive transport costs are quantitatively sizeable. On average over 1974-2013, the additive cost is estimated to be 1.8\% and 2.9\% of the export price in air and vessel transport respectively. This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period). Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport mode. To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs. This represents a valuable insight for calibrating related international trade or business cycle models. We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs. Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor~2.

@@

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian firm-level export data in 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of ad-valorem trade costs (expressed in Norwegian crowns). Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs. Yet, our paper complements their findings in two main respects. First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive to the ad-valorem cost; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs. From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect. While \citet{Irrazabal_2015} obtain that, for an import price of 100 USD, 6 USD are paid for additive trade costs, our own results point to a value of 2.7 USD paid in additive costs for maritime transport and 1.4 USD for air shipping. This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode). \footnote{See Section \ref{sec:results_decomposition} for more details.} Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} and \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

%\textbf{enleve ancienne footnote plus haut qui reprenait cette difference, a mettre en footnote ici? en commentaire en dessous}
%However, given their database and their empirical approach, \citet{Irrazabal_2015} can only identify the ratio of the additive to the ad-valorem cost ($\frac{t_{ik}}{\tau_{ik}}$ in our terminology), that they interpret by expressing it in terms of the median export price by country-product ($\bar{\widetilde{p}}_{ik}$). By contrast, our estimation strategy enables us to uncover both values of the ad-valorem and the additive costs $\tau_{ik}$ and $t_{ik}$ separately, in terms of the country-product export price $\widetilde{p}_{ik}$.

@@

@@

Under Model  \textit{(a)}, the estimated equation simplifies as:\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs. This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$. The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results. These results are reported in the Online Appendix, available on the authors' webpages.}

@@

After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component, $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode. When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is} = \widehat{\tau}_{i} \times \widehat{\tau}_{s}$. In this case, notice that the equation could be estimated relying on a linear form. To preserve comparability of the results, we keep the same non-linear estimation method in both cases though. Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme. We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{ice}$, \textit{b)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode. These results are reported in Section \ref{sec:results_decomposition}.

@@

\begin{table}[htbp]
  \centering
  \footnotesize{
  \caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    \multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
    Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
    \multicolumn{5}{l}{\textbf{Model (a) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice} -1$, in \%)}  \\ \hline
    Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
   \textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} & & & & \\ \hline
    Mean  & 3.2 & 2.5 & 3.3 & 2.4 \\
    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \textit{Additive term }& & & &   \\ \hline
    \textit{In \% of the export price} &&&& \\ \hline
        Mean  & 2.9 & 1.8 & 2.8 & 1.9 \\
    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
   \textit{ In USD per kg traded }&&&& \\ \hline
    Mean & 0.09	&1.12	&0.08&	2.16 \\
    Median & 0.07	&1.04	&0.06&	1.06 \\ \hline
    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
    \multicolumn{5}{l}{\textbf{Data}  } \\ \hline
    Transport costs ($p/\widetilde{p} -1$, in \%) & & & & \\ \hline
    Mean & 5.3 & 5.0& 5.6&3.9 \\
    Median & 4.3 & 2.0 & 4.4& 1.9 \\ \hline
    Export price ($\widetilde{p}$), in USD & & & & \\
    Mean & 16.0 &	6488.4	&9.6	&6643.6 \\
    Median & 4.1	& 142.5	& 4.1	& 142.2 \\ \hline
    \# obs. & 29279 & 28207 & 29317 & 27680 \\
    \# origin country & 188 & 191 & 188 & 189 \\
    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
  \end{tabular}
    \end{center}}
\parbox[l]{10cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value relative to total trade flows. The additive term is expressed in fraction of fas price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}\\
\parbox[l]{10cm}{\tiny{Notes: For the  4-digit classification, statistics for observed data have been calculated for the same set of years as used for estimation, i.e. 1974, 1981, 1989, 2001, 2009, 2013.}}
\end{table}%

@@

Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs. On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.12 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case. Expressed in terms of the export price, the additive costs amount to 2.9\% and 1.8 \% for ocean shipping and air transport respectively. This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\% and 1.8\% for vessel and air shipping respectively (on average over the period). Put it differently, the additive component amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air. This shows that the per-kg cost dimension is quantitatively sizeable. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly a factor 2. Additionally, we run quality-of-fit diagnostic tests that confirms the importance of the additive cost component. The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degree of freedom, as reported in Appendix \ref{app:diagnostic_test}. Further, these results provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration.\smallskip

Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question. Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of ad-valorem trade costs (expressed in Norwegian crowns), ie $\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and picking up their estimated weighted mean value similarly to us). One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem). From this, we can rebuilt the ratio in similar terms to them -- allowing us to gain in generality in this respect. Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.4\% in air transport. This may sound surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}. This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case. While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage). In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment. This obeys the following reasoning. \citet{Irrazabal_2015}'s results imply that, for a import price of 100 USD, 6 USD are paid in additive trade costs. For the same import price of 100 USD, our results yield the value of 2.7 (1.4) USD paid in additive transport costs for maritime (air) transport.\footnote{In both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$. With $x = 0.028$ and $0.014$ in vessel and air transport  in 2004, we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 0.027 and 0.014 in each transport mode respectively.} This comparison of results suggests that between 25\% (in Air) and 48\% (in Vessel) of additive trade costs are attributable to international transport. This, on top of our results summarized in Table \ref{tab:summary_results}, point out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs.

\textbf{stop here. Grosse interrogation sur notre positionnement par rapport a Hummels dans la section suivante. Ai l'impresssion qu'on explique mal nos differences de resultats. On a l'air de dire que, contrairement a Hummels, les transport costs per se ont bien baisse, et que c'est ca qui compte. En regardant la figure "Characterizing the time trends: Applying Hummel's (2007) method", il me semble que c'est mal dit. Hummels ne dit pas que les couts purs n'ont pas baisse, d'ailleurs il montre qu'ils ont baisse, mais qu'au sein d'un mode de transport donne, ils sont en partie compense par les effets de composition. Nous, on montre que les couts purs ont baisse, et que les effets de composition, soit ne jouent pas (air), soit amplifient (au lieu de partiellement compenser, vessel) la baisse des couts purs. Il me semble qu'il faut reecrire l'interpretation des resultats en ce sens.}

@@

\paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}. He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f$ the shipping charge per kg shipped. Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, the transport cost measure writes down as:

@@

We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs. To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price, ie $\beta$, according to:
\begin{eqnarray*}
\beta_{ikt} &\equiv& \frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}-1}
\end{eqnarray*}

@@

The above results point out that the share of additive costs in total transport costs is not constant across the triplet year/product/origin country, thereby contrasting the implicit Hummel's (2007) assumption. This drives us to provide a decomposition of the transport costs trends (between the trade composition effects and the ``pure'' transport costs time trends), that explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} An important contribution of the paper is to show the importance of this (empirically relevant) in analyzing the underlying determinants of transport costs trend patterns. We start presenting our estimation strategy.\smallskip
